---
title: "Introduction to R programming"
author: "Rita Giordano"
date: "giordanor@rsc.org, rgiordano@gmx.com"
output:
  ioslides_presentation:
    #logo: RSC_logoS.png
  beamer_presentation: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
```

<script src="https://ajax.googleapis.com/ajax/libs/jquery/1.12.2/jquery.min.js"></script>

<script>
    $(document).ready(function() {
      $('slide:not(.title-slide, .backdrop, .segue)').append('<footer label=\"Introduction to R programming ECM32 
      Crystallographic Computing Forum Melk 14th-17th August 2019\"></footer>');    
    })
</script>

<style>
footer:after {
    content: attr(label);
    font-size: 12pt;
    position: absolute;
    bottom: 20px;
    left: 100px;
    line-height: 1.9;
  }
  .gdbar img {
  width: 300px !important;
  height: 150px !important;
  margin: 8px 8px;
}

.gdbar {
  width: 400px !important;
  height: 170px !important;
}
  
</style>

<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 8px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 28px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>


## R introduction
R is a programming language for statistical computing.

Useful website:

- https://www.r-project.org
- https://cran.r-project.org
- https://www.r-bloggers.com
- https://en.wikibooks.org/wiki/R_Programming

## Packages

- crone: https://cran.r-project.org/web/packages/crone/index.html
- cRy (J. Foadi, D. Waterman, R.Giordano): to analyse mtz file and visualize data ...work in progress!!!
- bio3d   protein structure analysis https://thegrantlab.org/bio3d/index.php
- tidyverse 
- ggplot2 - grammar for graphics

## RStudio

RStudio is an IDE for R: 

- https://www.rstudio.com

- https://www.rstudio.com/products/rstudio/download/#download

## How to assign values to a variable and load package

In R to assign a value to a variable there is this symbol `<-`

called assignment operator.

```{r cars, echo = TRUE, eval=TRUE}
x <- 5
x
```

To load package use the function library
```{r, echo = TRUE, eval=TRUE}
library(tidyverse)
library(cry)
```

## Control flow past and present

Before it was:

- if-else: 
```{r,echo = TRUE,eval=FALSE}
if (condition) {expression}
else
```

Now if_else {dplyr} a vectorised if:

```{r,echo = TRUE,eval=FALSE}
if_else(condition, true, false)
```
- for:
```{r,echo = TRUE,eval=FALSE}
for (var in seq) {expression}
```


## Control flow

- repeat:
```{r,echo = TRUE, eval=FALSE}
repeat {expression}
```
- break: to stop loop
- next: to skip iteration
- return: to exit from a loop and return a value

## R function

An R function is defined as global environment: .GlobalEnv.
```{r,echo = TRUE, eval=TRUE}
add <- function(x,y)
{
  a = x+y  
  
  return(a)

}

add(2,3)
```

```{r, echo = TRUE, eval=TRUE}
environment(add)
```



## Tidyverse


“Happy families are all alike; every unhappy family is unhappy in its own way.” –– Leo Tolstoy

“Tidy datasets are all alike, but every messy dataset is messy in its own way.” –– Hadley Wickham



## Read csv_file

```{r, echo = TRUE, eval=TRUE}
spot <- read_csv("XRAy_files/spot.csv")
```

## Create summary statistics

```{r, echo = TRUE, eval=TRUE}
summary(spot)
```


## Check data frame and columns names.
```{r, echo = TRUE, eval=TRUE}
head(spot)
names(spot)
```


## `$`
The simbol `$` can access to the columns of a data frame.
```{r, echo = TRUE, eval=TRUE}
head(spot$Peak)

class(spot)
```

## Rename columns
Use pipe
```{r, echo = TRUE, eval=TRUE}
spot <- spot %>%
          rename(id = X1,
                 X = Coor.1,
                 Y = Coor.2,
                 Intensity = Peak)
head(spot)
```


## Filter values of a column

```{r, echo = TRUE, eval=TRUE}
spot <- spot %>%
          filter(Intensity > 1000)
head(spot)
```

## Sort data

```{r, echo = TRUE, eval=TRUE}
spot <- spot %>%
          arrange(Intensity)
head(spot)
```

## Select column and create new variable

```{r, echo = TRUE, eval=TRUE}
spot <- spot %>%
          select(-c('MeasB.1', 'MeasB.2', 'MeasB.3', 'MeasB.4')) %>%
          mutate(code = if_else(Intensity > 25000, 'strong', 'medium'))
head(spot)
```

## Plot Intensity using ggplot2

```{r, echo = TRUE, eval=TRUE}
spot %>%
  ggplot(aes(X, Y, size = Intensity, color = code)) +
    geom_point() + theme_bw()

```